---
title: "RCDS Introduction to sampling & hypothesis testing"
author: "Jesus Urtasun; jurtasun@ic.ac.uk"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::robobook:
    self_contained: true
    code_folding: hide
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Chapter 2. Confidence intervals and the central limit theorem

## Import libraries

```{r import, include = FALSE}

# Import libraries

```

## 2.1 The central limit theorem

A random event is an event whose output we don't know. For addressing such events, we define a quantity we will call probability, a number 𝑝 between 0 and 1 quantifying the degree of uncertainty or "surprise" associated to the output of a measurement.

### i) Consider a random variable

```{r binomial, include = FALSE}

# Consider a random variable

```

### ii) Small sample size

```{r small_n, include = FALSE}

# Small sample size

```

### iii) Large sample size

```{r large_n, include = FALSE}

# Large sample size

```
