---
title: "Introduction to random sampling and hypothesis testing"
author: "Research Computing and Data Science: Jesus Urtasun; jurtasun@ic.ac.uk"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::robobook:
    self_contained: true
    code_folding: hide
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r import, cache = T, echo = T, eval = T, warnings = F, messages = F}

# Import libraries

# Set working directory
workdir <- "/Users/jurtasun/Desktop/Courses/ICL/RCDS_Introduction_sampling_hypothesis_testing/Chapter2"
setwd(workdir)

set_plot_dimensions <- function(width_choice, height_choice) {
    options(repr.plot.width = width_choice, repr.plot.height = height_choice)
}

```

# Chapter 2

## 2.1 The Central Limit Theorem

In this chapter we will describe one of the most fundamental results in statistical inference. The central limit theorem (CLT) states that the distribution of sample means, meaning the mean values computed for a set of different experiments, approximates a normal distribution as the sample size gets larger, regardless of the population's distribution. Sample sizes equal to or greater than 30 are often considered sufficient for the CLT to hold.

### Event binomial distributed

```{r clt1, cache = T, echo = T, eval = T, warnings = F, messages = F}

# Binomial distribution
n <- 10
p <- 1/6
x <- 0:n
     
# Probability mass function
pmf <- dbinom(x,n,p)  # a binomial distribution with n=10, p=1/6

# Plot probability mass function
set_plot_dimensions(5, 4)
plot(x, pmf, ylim = c(0, 0.4), xlab = "x", type = "h", col = "red", axes = FALSE)
points(x, pmf, col = "red")
axis(side = 1, at = 0:10)
axis(side = 2)

# Compute expected value
expectation <- sum(x * pmf)
expectation

# Compute variance
variance <- sum((x - expectation)^2 * pmf)
variance

```

### Simulate 5 repetitions

```{r clt2, cache = T, echo = T, eval = T, warnings = F, messages = F}



```
