---
title: "RCDS Introduction to sampling & hypothesis testing"
author: "Jesus Urtasun; jurtasun@ic.ac.uk"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::robobook:
    self_contained: true
    code_folding: hide
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Chapter 2. Confidence intervals and the central limit theorem

## Import libraries

```{r import, include = FALSE}

# Import libraries

```

## 2.2 Confidence intervals

We just described the central limit theorem and how the distribution of sample means approximates a normal distribution as the sample size gets larger. Sample sizes equal to or greater than 30 are often considered sufficient for the CLT to hold.

We will now discuss how to building confidence intervals, meaning a range in which we know a certain value will fall. As a first example, we will work with the standard normal distribution and "standardized" variables. The best single estimate of the population mean $\mu$ based on a sample of data is simply the sample mean, $\bar{x}$.

### i) Standard normal distribution

```{r standard, include = FALSE}

# Standard normal distribution

```

### ii) Z-score and standard variables

```{r z_score, include = FALSE}

# Z-score and standard variables

```

### iii) Confidence intervals for large N

```{r large_n, include = FALSE}

# Confidence intervals for large N

```

### iii) Confidence intervals for small N

```{r small_n, include = FALSE}

# Confidence intervals

```