---
title: "RCDS Introduction to sampling & hypothesis testing"
author: "Jesus Urtasun; jurtasun@ic.ac.uk"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::robobook:
    self_contained: true
    code_folding: hide
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Chapter 3. Hypothesis testing

## Import libraries

```{r import, include = FALSE}

# Import libraries

```

## 3.1 General approach

There are many circumstances where we simply want to check whether an observation looks like it is compatible with the null hypothesis, 𝐻0. Having decided on a significance level $\alpha$ and whether the situation warrants a one-tailed or a two-tailed test, we can use the cdf of the null distribution to calculate a p-value for the observation.

Parametric tests rely on a probability distribution of known form as a model for the null hypothesis. Here we will look at some commonly-encountered examples. How surprising is my result? Calculating a p-value

### i) Example: check if a die is loaded

Your arch-nemesis Blofeld always seems to win at ludo, and you have started to suspect him of using a loaded die.
You observe the following outcomes from 100 rolls of his die:

```{r example1, include = FALSE}

# Check if a die is loaded

```

### ii) Example: check if a coin is fair

Dr Vogel has challenged you to a game of ludo and you have agreed to flip a coin to decide who starts.
You're not sure whether the coin she is using is fair or not.
She flips it 50 times for you, with the following results: (1=heads, 0=tails)

```{r example2, include = FALSE}

# Check if a coin is fair

```

### 3.2 Comparing means: 2-sample t-test

```{r t_test, include = FALSE}

# Comparing means: 2-sample t-test

```
