---
title: "RCDS Introduction to sampling & hypothesis testing"
author: "Jesus Urtasun; jurtasun@ic.ac.uk"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::robobook:
    self_contained: true
    code_folding: hide
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Chapter 1. Random variables and probability distributions

## Import libraries

```{r import, include = FALSE}

# Import libraries

```

## 1.1 Random events and probability

A random event is an event whose output we don't know. For addressing such events, we define a quantity we will call probability, a number 𝑝 between 0 and 1 quantifying the degree of uncertainty or "surprise" associated to the output of a measurement.

```{r probability, include = FALSE}

# What is probability

```

## 1.2 Discrete distributions

Discrete distributions, also referred to as "probability mass distributions", address random events where the number of possible outcomes is a finite, or "countable" number. Processes such as tossing coins, rolling dice, or anything related to counting, fall in this category.

### i) Binomial distribution

```{r binomial, include = FALSE}

# Binomial distribution

```

### ii) Poisson distribution

```{r poisson, include = FALSE}

# Poisson distribution

```

## 1.3 Continous distributions

In the case of continuous variables, such as measuring most physical quantities, the output of an experiment or measurement is typically a continuous, or "uncountable" number, lying in an infinite interval. For such processes, the intuitive definition of probability we discussed previously, needs to be replaced by a new mathematical object, that we will call continuous distribution, or "probability density distribution".

### i) Uniform distribution

```{r uniform, include = FALSE}

# Uniform distribution

```

### ii) Exponential distribution

```{r exponential, include = FALSE}

# Exponential distribution

```

#### iii) Gaussian distribution

```{r gaussian, include = FALSE}

# Gaussian distribution

```